A Parallel General Purpose Multi-Objective Optimization Framework, with Application To Electron Beam Dynamics

### 2019-01-31 paper/submission/prab.tex
incorporating Linda's edits to abstract

\begin{abstract}
Particle accelerators are invaluable tools for research in the basic and
  applied sciences, such as materials science, chemistry,
  the biosciences, particle physics, nuclear physics and medicine.
The design, commissioning, and operation of accelerator facilities is a
  non-trivial task, due to the large number of control parameters and the
  complex interplay of several conflicting design goals.
We propose to tackle this problem by means of multi-objective optimization
  algorithms which also facilitate a parallel deployment.
In order to compute solutions in a meaningful time frame, a fast 
  and scalable software framework is required.
In this paper, we
present a general-purpose framework for simulation-based
multi-objective optimization methods that allows the automatic
investigation of optimal sets of machine parameters.
Using evolutionary algorithms as the optimizer and \textsc{OPAL} as the 
  forward solver, validation experiments and 
 results of multi-objective optimization problems in the domain of beam
  dynamics are presented. The high charge beam line at the Argonne Wakefield 
  Accelerator Facility was used as the beam dynamics model. 
  Optimized solutions found by the framework were used to finish the design 
  of a two beam acceleration experiment.

>>>>

\begin{abstract}
Particle accelerators are invaluable tools for research in the basic and applied sciences, such as materials science, chemistry,
  the biosciences, particle physics, nuclear physics and medicine. The design, commissioning, and operation of accelerator facilities is a
  non-trivial task, due to the large number of control parameters and the complex interplay of several conflicting design goals.
  The Argonne Wakefield Accelerator facility has some unique challenges resulting from its purpose to carry out advanced accelerator R\&D.
  Individual experiments often have challenging beam requirements, and the physical configuration of the beamlines is often changed
  to accommodate the variety of supported experiments. The need for rapid deployment of different operational settings
  further complicates the optimization work that must be done for multiple constraints and challenging operational regimes. 
  One example of this is an independently staged two-beam acceleration experiment which requires the construction 
  of an additional beamline (this is now in progress).  The high charge drive beam, well into the space charge regime, must be threaded
  through small aperture (\SI{17.6}{mm}) decelerating structures.  
  In addition, the bunch length must be sufficiently short to maximize power generation in the decelerator.  
We propose to tackle this problem by means of multi-objective optimization algorithms which also facilitate a parallel deployment.
In order to compute solutions in a meaningful time frame, a fast and scalable software framework is required.
In this paper, we present a general-purpose framework for simulation-based
multi-objective optimization methods that allows the automatic investigation of optimal sets of machine parameters.
Using evolutionary algorithms as the optimizer and \textsc{OPAL} as the forward solver, validation experiments and 
 results of multi-objective optimization problems in the domain of beam dynamics are presented. 
  Optimized solutions for the new high charge drive beamline found by the framework were used to finish the design 
  of a two beam acceleration experiment.
  The selected solution along with the associated beam parameters is presented.
